Initial Analysis

### 2019-05-24 Initial_Analysis.Rmd
Ready for first try in THE BEAST

```{r Spatial_Data, eval = T, echo = T, results='hide'}

# The path
path_world <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Spatial_Data/SAU_Shapefile/"
#The File
fnam_world <- "SAUEEZ_July2015.shp"
#Load it!

# World_EEZs_sf <- st_read(dsn = path_world,
#                       layer = file_path_sans_ext(fnam_world))

# head(World_EEZs)

EEZIDs_List <- read_excel("~/Documents/UBC/Oceans_Project/Distribution/By Cell/Cell_Data/Updated_EEZList_17June2016.xlsx")
EEZ_CellID <- read_excel("~/Documents/UBC/Oceans_Project/Distribution/By Cell/Cell_Data/EEZ_CellID.xlsx")

# DBEM Coordinate system
Coor <- fread("~/Documents/UBC/Oceans_Project/Distribution/DBEM/Data/Lon_Lat_DBEM.txt",header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

#### Estimate the number of grid-cells that each EEZ as

EEZ_Size <- EEZ_CellID %>% 
  group_by(EEZID) %>% 
  summarise(n_tot_grid = n()) %>% 
  left_join(EEZIDs_List,
            by = "EEZID")

@@

  Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/"
  INDEX <- seq(1,259200,1)

@@

    # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
    SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
      select(CATCH.56:CATCH.65) %>% 
      mutate(INDEX = INDEX) %>% 
      gather("Year","Catch",1:10) %>% # Last 10 years of data
      group_by(INDEX) %>% 
      summarise(SAU_C = mean(Catch,na.rm=T))

    # Join both tables
    sppdist <- sppdist %>% 
      left_join(SAU_C_data,
                by = "INDEX") %>% 
      select(TaxonKey,INDEX,everything())

@@

@@

# The function
EstTransIndex <- function(Spp,Model,EEZ_Size,Neighbors){

  # Get model data from spps
  SppDist <- GetSppDist(600004,"All")

  Index_EEZ <- Neighbors %>% 
    select(INDEX)
      # Result 1. Number of Countries that share the species
  # Trans_Index

  Trans_Spp <- suppressWarnings(SppDist %>%
    gather("Model","Value",3:6) %>%
    filter(
      INDEX %in% Index_EEZ$INDEX,
      Value != 0
      ) %>%
    left_join(Neighbors,
              by = "INDEX") %>% # Match the species distribution to the Countries EEZs
    group_by(Name = Country_Territory,
             Neighbor_Territory,
             TaxonKey,
             Model
             ) %>% 
    summarise(n_cells_spp = n()) %>% # The number of cells present within each country's EEZ
    left_join(EEZ_Size,
              by = "Name") %>% 
    mutate(Trans_Index = n_cells_spp/n_tot_grid) %>%  # Include one transboundary species
    select(EEZID,everything())
    ) # For the sake of my brain

  EstTransIndex=Trans_Spp # A dataframe with each country and neighbouring states sharing the species

} # End function

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# EEZ_Size <- EEZ_Size
# Neighbors <- Neighbor_List

@@

@@

# Set Spceies list
Exploited_Species <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv") %>% 
  filter(TaxonKey %in% Species_List)

@@

Test_Countries <-c("Guatemala (Caribbean)","Guatemala (Pacific)","Belize","Honduras","El Salvador","Costa Rica (Pacific)","Costa Rica (Caribbean)", "Panama (Caribbean)", "Panama (Pacific)")

Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

@@

# Exploited_Species <- c("600004","600006",600107)
# Exploited_Species = 600004
# TaxonKey <- unique(Spps$TaxonKey)

@@

@@

Spp_Trans <- NULL

system.time(
  Spp_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List, FUN=EstTransIndex, EEZ_Size = EEZ_Size, Model = "All", Neighbors = Neighbor_List, mc.cores = n_Cores
    )
  )
)

>>>>

@@

  Distpath <- paste(Data_Path,"Distribution_Data/",sep="")
  INDEX <- seq(1,259200,1)

@@

    # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
    SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
      select(CATCH.1:CATCH.65) %>% 
      mutate(INDEX = INDEX) %>% 
      gather("Year","Catch",1:10) %>% # Last 10 years of data
      group_by(INDEX) %>% 
      summarise(SAU_C = mean(Catch,na.rm=T))

    # Join both tables
    sppdist <- sppdist %>% 
      left_join(SAU_C_data,
                by = "INDEX") %>% 
      select(TaxonKey,INDEX,everything()) %>% 
      gather("Model","Value",3:6) %>%
      mutate(Value = ifelse(is.na(Value), 0, Value))# Converting NA's to ceros

@@

#### NOTES:

##### General:
- For now using SAU_C as presence absence
- Grouping by species distribution within EEZ and not EEZ area

##### Tresholds:
##### Model Index, line ~479

This first treshold estimates the uncertanty of models. Goes from 0 - 1 as follows:

- 1 or 0; All models agree that in the presence of the species
- 0.75; Three out of four models agree on speices presence
- 0.50; Two out of four models agree on speices presence
- 0.25; Only one model flags a speices presence
- 0; All models agree there is no species there

##### Distribution Index, line ~482

@@

# The function
EstTransIndex <- function(Spp,Model,Neighbors){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model)

  # Get the index for testing
  Index_EEZ <- Neighbors %>% 
    select(INDEX)

  # Estimate total spp distribution within neighbours
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Index_EEZ$INDEX) %>% 
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)) %>% 
    filter(Model_Index > 0) %>% 
    mutate(Model_Index = Model_Index*100) %>% 
    left_join(Neighbors,
              by = "INDEX")

  # Estimating total grids per country
  Spp_Grid <- Trans_Spp %>% 
    group_by(Name = Country_Territory,
             Neighbor_Territory,
             Model_Index) %>% 
    summarise(n_cells_spp = n()) %>% 
    ungroup()

  # Split dataframes to merge latter
  Tone <- select(Spp_Grid,1,3,4)
  Ttwo <- select(Spp_Grid,Name=2,3,4) # Change column name

  # Merge dataframes
  Merged_data <- full_join(Tone,Ttwo, by = c("Name","Model_Index")) %>% 
    rowwise() %>% 
    mutate(Spp_Total = sum(n_cells_spp.x,n_cells_spp.y,na.rm=T)) %>% 
    select(Name,Model_Index,Spp_Total) %>% 
    left_join(Spp_Grid,
              by = c("Name","Model_Index")) %>% 
    ungroup() %>% 
    group_by(Name,Model_Index) %>% 
    summarise(Spp_Total = sum(Spp_Total))

  # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Indexes <- Trans_Spp %>% 
    group_by(Name,
             Neighbor_Territory,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% # The number of cells present within each country's EEZ
    #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 2)_________ #
    left_join(Merged_data,
              by = c("Name","Model_Index")) %>% 
    mutate(Trans_Index = n_cells_spp/Spp_Total)

  ### Save spp dataframe

  if(nrow(Indexes) > 0){

    File_Name <- paste(Spp,"_Transboundary.csv",sep = "")
    Save_Path <- paste(Results_Path,File_Name,sep="")

    write.csv(Indexes,
              Save_Path,
              row.names = F)

  ### Function return

  EstTransIndex=Indexes # A dataframe with

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# Neighbors <- Neighbor_List

# Spp <- NULL
# Model <- NULL

@@

#### Set paths for computer ####

# Set paths depending on machine Beas or Hall1000
if(Sys.info()[7] == "jepa88"){
  Data_Path <- "Z:/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/",sep = "")
}else{
  Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/"
  Path <- paste(Data_Path,"Results/",sep = "")
}

# Path for saving Results

# If path exists overrite it, otherwise creats a new path for that year
if(file.exists(Path) == TRUE){
  Results_Path = Path
}else{
  dir.create(Path)
  Results_Path = Path
}

#### Load Data Needed ####

# The path
path_world <- paste(Data_Path,"Spatial_Data/SAU_Shapefile/",sep="")
#The File
fnam_world <- "SAUEEZ_July2015.shp"
#Load it!

EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))

# DBEM Coordinate system
Coor <- fread(paste(Data_Path,"Spatial_Data/Lon_Lat_DBEM.txt",sep=""),header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

@@

# Set Spceies list
Exploited_Species <- fread(paste(Data_Path,"Distribution_Data/exploited_species_list.csv",sep="")) %>% 
  filter(TaxonKey %in% Species_List)

@@

# Test_Countries <-c("Guatemala (Caribbean)","Guatemala (Pacific)","Belize","Honduras (Pacific)","El Salvador","Costa Rica (Pacific)","Costa Rica (Caribbean)", "Panama (Caribbean)", "Panama (Pacific)")

# Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

@@

# Exploited_Species_List <- c(600004,"600006",600107)
# Exploited_Species_List = 600006
# TaxonKey <- unique(Spps$TaxonKey)

@@

# Step 1; EstNeighbordList 
Neighbor_List <- EstNeighbors(World_EEZs)
# head(Neighbor_List)

@@

# Spp_Trans <- NULL

system.time(
  Spp_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List[1], FUN=EstTransIndex, Model = "All", Neighbors = Neighbor_List, mc.cores = n_Cores
    )
  )
)
